\documentclass{book}
\usepackage[T1]{fontenc}
%\usepackage[standard-baselineskips]{cmbright}
\usepackage{cmbright}
\def\usedfonts{CM-Bright}
\usepackage{graphicx} % includegraphics
\begin{document}
%=========================================================================
\section*{Status}
%=========================================================================



%=========================================================================
\section*{Overview}
%=========================================================================
The benchmark Element0ktest is based on the work by Lejaeghere
\cite{lejaeghere14_critrevsolstmatsci39_1,lejaeghere13_arxiv1204_2733}.
It includes the Elemental structures at zero Kelvin. The equation of
state, i.e. equilibrium volume, bulk modulus and the derivative of the
latter are compared the data calculated with the Wien2k code.

The test consists of
\begin{center}
\begin{tabular}{|l|l|}
\hline
bin/Strcfiles & directory holding all structure files, yet without species
information \\
bin/el\_makedo & bash script which rebuilds the input data in the Cases directory \\
bin/specieslist & container of the species information for all elements\\
bin/splitspecies.f90 & code to split up specieslist\\
\hline
\end{tabular}
\end{center}

To build up the test, 
\begin{enumerate}
\item change the file specieslist so that the desired setup
  construction is selected for each element.
%
\item construct an bash script to run the jobs
\begin{verbatim}
bin/el_makedo
\end{verbatim}
This creates a new bash script el\_do\_sample. It also updates the
setup information: The file \verb|src/specieslist| is expanded into
the files in the directory \verb|bin/Speciesfiles|.
%
\item copy \verb|el_do_sample| into \verb|el_do|. Unselect elements
  that shall not be studied by deleting or by commenting out the
  corresponding lines in the list.
%
\item construct a file \verb|sample.cntl|.
\begin{verbatim}
cp src/sample.cntl_strt sample.cntl
\end{verbatim}
This control file which will be used by all calculations. Take special
care to remove \verb|START=T|, when the restart files are present and
if they shall be used.
%
\item perform the paw calculation by executing 
\begin{verbatim}
el_do -j 3 -- 1>out 2>&1 &
\end{verbatim}
The calculation may take up to a few days. The calculations fill up
the directory \verb|Cases| which will contain one subdirectory for
each element.
%
\item when the calculations are finished, update the result files \verb|etot.dat|, \verb|gap.dat| and \verb|murn.in| with
\begin{verbatim}
el_do -u
\end{verbatim}

\end{enumerate}

%=========================================================================
\section*{Experiences}
%=========================================================================
For transition metals we need to set !POT:VAL0=0. 

For Pd we had to increase  !POT:POW from 3. to 5.



\end{document}
